\section{Discussion}
\label{sec:discussion}

\subsection{What multi-seed replication revealed}

The three-seed replication substantially changes the picture relative to the single-seed results, in two important ways.

\textbf{Single-seed dev was misleading.} Seed 42 happened to be a favorable initialisation for Norwegian XLM-R (86.49 vs 85.52). Across three seeds, English XLM-R averages marginally higher on dev (85.87 vs 85.45). Had we reported only seed 42 and drawn conclusions from dev performance, we would have overstated the Norwegian advantage.

\textbf{The test set tells a clearer story.} On the held-out test set, Norwegian is consistently ahead for both encoders: $+$1.86 F1 for XLM-R ($p=0.10$, borderline) and $+$3.36 F1 for mBERT ($p=0.046$, significant). The mBERT result is the most reliable: the gap is large, variance is low for both systems, and statistical significance is achieved even with only two usable seed pairs. The XLM-R result is directionally consistent but remains borderline; a fourth or fifth seed would likely tip it to significance.

The honest summary: Norwegian is a significantly better zero-shot transfer source than English for mBERT on the Danish test set, and a consistently (but not yet significantly) better source for XLM-R. The case is strongest on test data and weakest on dev, which reinforces the standard recommendation to report held-out test performance rather than dev performance for cross-lingual comparisons.

\subsection{Inconvenient finding: EN mBERT best fine-tuned, DA XLM-R$<$DA BERT}

Two results in Table~\ref{tab:dev_results} cut against the paper's preferred narrative and deserve direct attention.

\textbf{English mBERT fine-tuned on Danish is the best transfer-plus-fine-tuning model in the whole study (89.4 dev, 85.5 test)}, outperforming every Norwegian transfer variant. This directly contradicts a naive reading of the linguistic-distance hypothesis: if proximity predicts transfer quality, the Norwegian fine-tuned models should rank highest. The likely explanation is that mBERT trained on English benefits from \emph{more English training data} (English has 20,142 sentences vs Norwegian's 4,183), which gives it a richer representation before fine-tuning, and that this data-volume advantage outweighs the typological-distance advantage of Norwegian. It also reflects the fact that EN→DA mBERT FT starts from a checkpoint trained on five times as much labeled data. The distance hypothesis operates most cleanly in the \emph{zero-shot} regime, where data volume cannot compensate. Once fine-tuning on Danish is introduced, the data-quality and initialisation effects dominate.

\textbf{Danish XLM-R underperforms Danish BERT on Danish-only training} (88.6 vs 89.0 dev, 84.9 vs 85.3 test). This is unexpected given XLM-R's generally stronger cross-lingual results. A plausible explanation is that \texttt{xlm-roberta-base} allocates vocabulary and attention across 100 languages and may be slightly undertrained on Danish relative to \texttt{bert-base-multilingual-cased}, which uses a shared vocabulary more efficiently for smaller language sets. Another possibility is that the difference is within run-to-run noise, and a multi-seed experiment would show no consistent gap. Either way, the result weakens the blanket claim that XLM-R is the strongest encoder.

\subsection{Encoder choice}

XLM-R leads on source-language baselines (92.0 vs 91.2 for Norwegian, 81.7 vs 80.1 for English) and on zero-shot transfer (86.5 vs 83.4 for Norwegian zero-shot). The XLM-R advantage is most consistent in zero-shot settings; it narrows or reverses after fine-tuning on Danish. This suggests that XLM-R's broader coverage helps when the model must generalise without target-language data, but mBERT's more concentrated vocabulary can be competitive once Danish labels are available.

\subsection{Data efficiency}

The Danish learning curve (Figure~\ref{fig:learning_curve}) shows that transfer is useful even before any Danish supervision is added, but a modest amount of target-language annotation quickly recovers most of the full-data performance. The Danish XLM-R model reaches the English zero-shot reference at around 50\% of the training data and approaches the Norwegian zero-shot reference at 50--100\%. In practical terms: for a project with a limited annotation budget, starting with a Norwegian zero-shot model and labeling $\sim$50\% of the Danish training set is a reasonable cost-effective strategy.

\subsection{Limitations}
\label{sec:limitations}

The main limitations are: (1) single-seed experiments, meaning effect sizes are uncertain; (2) only English and Norwegian as source languages, while Swedish or German experiments would strengthen the distance-gradient claim; (3) evaluation on one NER benchmark family, so results may not generalise to other Danish NER datasets or domain-shifted text; (4) no hyperparameter tuning per source language, which may disadvantage some systems.
