\section{Introduction}

Named entity recognition (NER) identifies text spans referring to persons, locations, and organizations. For Danish, the bottleneck is data rather than model capacity: English offers abundant labeled resources; Danish does not \citep{hvingelby2020dane}. Cross-lingual transfer addresses this by training on a well-resourced language and deploying to the target with no labels, or before any exist.

The key question is which source language best serves Danish. Norwegian is the natural candidate: both are North Germanic languages sharing substantial syntax, morphology, and vocabulary. English is the conventional high-resource default. We quantify this contrast using \texttt{lang2vec} WALS syntactic feature vectors \citep{littell2017uriel}: Norwegian is 5.7$\times$ closer to Danish than English (cosine distance 0.021 vs.\ 0.119), with Swedish intermediate at 0.039 and German more distant at 0.124. Swedish and German serve as reference points confirming the coherence of the North Germanic cluster and anchoring the broader Germanic comparison. If structural similarity shapes representation alignment inside multilingual encoders, Norwegian-trained models should transfer more effectively. Transfer from English to Danish has been shown to work \citep{plank2020danish}; our contribution is a controlled Norwegian comparison with explicit typological grounding, holding encoder family, data format, and evaluation protocol constant.

We run five experiment families: monolingual source baselines, zero-shot transfer to Danish, transfer plus Danish fine-tuning, Danish-only supervision at varying data budgets, and a typological-distance analysis. Encoder family is fixed (mBERT and XLM-R) across all conditions so that performance differences are attributable to source language. A secondary aim is to disentangle three factors that are typically conflated: source-language choice, encoder architecture, and the amount of Danish supervision available. All comparisons span three random seeds with paired resampling tests.

\paragraph{Main research question.}
\textit{Does lower typological distance to Danish predict better NER transfer: is Norwegian a more effective zero-shot source than English?}

\noindent\textbf{RQ1.} How do monolingual baselines compare to cross-lingual transfer routes?

\noindent\textbf{RQ2.} What is lost in zero-shot transfer, and how much does Danish fine-tuning recover?
