\section{Linguistic Distance Analysis}

\subsection{Motivation}

The proposal argued that Norwegian should transfer better than English to Danish because it is linguistically closer. This section quantifies that claim with a concrete, reproducible distance measure, then tests whether the observed NER transfer gap is consistent with the measured distance.

\subsection{Method}

We use \texttt{lang2vec} \citep{littell2017uriel}, which provides language feature vectors derived from the World Atlas of Language Structures (WALS). We use the \texttt{syntax\_wals} feature set, which encodes 103 typological features covering word order, morphological type, and related syntactic properties. Features not available for a given language pair are excluded before computing cosine distance; the distance between $L_1$ and $L_2$ is
\[
d(L_1, L_2) = 1 - \frac{\mathbf{v}_1 \cdot \mathbf{v}_2}{\|\mathbf{v}_1\|\,\|\mathbf{v}_2\|},
\]
where $\mathbf{v}_i$ contains only the features available for both languages.

\subsection{Results}

Table~\ref{tab:language_distance} reports pairwise cosine distances for English, Danish, Norwegian (Bokmål), Swedish, and German. Swedish and German are included as additional reference points: Swedish is the closest North Germanic neighbor of Danish after Norwegian, while German is a West Germanic language at greater distance.

\begin{table}[t]
\centering
\small
\resizebox{\linewidth}{!}{%
\begin{tabular}{lccccc}
\toprule
          & English & Danish & Norwegian & Swedish & German \\
\midrule
English   & 0.000   & 0.119  & 0.109     & 0.080   & 0.145  \\
Danish    & 0.119   & 0.000  & \textbf{0.021}     & 0.039   & 0.124  \\
Norwegian & 0.109   & 0.021  & 0.000     & 0.065   & 0.102  \\
Swedish   & 0.080   & 0.039  & 0.065     & 0.000   & 0.164  \\
German    & 0.145   & 0.124  & 0.102     & 0.164   & 0.000  \\
\bottomrule
\end{tabular}%
}
\caption{Pairwise cosine distances based on WALS syntactic features (lang2vec \texttt{syntax\_wals}). Lower = more similar. Norwegian is the closest language to Danish (0.021), followed by Swedish (0.039); English and German are substantially more distant ($\approx$0.12).}
\label{tab:language_distance}
\end{table}

Key observations:
\begin{itemize}
  \item Norwegian is the closest language to Danish (0.021), 5.7$\times$ closer than English (0.119).
  \item Swedish is the second closest (0.039), confirming the North Germanic cluster.
  \item English and German are at comparable distances (0.119 and 0.124), placing them both in a clearly more distant tier.
  \item The gradient Norwegian $<$ Swedish $<$ English $\approx$ German $<$ \dots\ is consistent with historical and genealogical accounts of Germanic language relationships.
\end{itemize}

\subsection{Interpretation}

With four reference languages the distance ordering is not fit to a single data point but reflects a consistent pattern. If linguistic distance predicts transfer quality, we expect: NO$\rightarrow$DA $>$ SV$\rightarrow$DA $>$ EN$\rightarrow$DA $\approx$ DE$\rightarrow$DA in zero-shot settings. We do not have Swedish and German transfer experiments, but the NO$\rightarrow$DA $>$ EN$\rightarrow$DA ordering is directionally confirmed by our results, even though the gap is not statistically significant under bootstrap testing (see Section~\ref{sec:results}).

It is important not to over-interpret this analysis. Cosine distance over WALS features is a proxy for structural similarity, not a causal mechanism. The same multilingual encoder processes all four source languages, and encoder-internal effects (training data volume, tokenizer overlap) may be as important as typological distance. The distance numbers strengthen the empirical story but do not prove that structure is the causal driver.
