\section{Methodology}

\subsection{Experimental design}

We evaluate five families of experiments:
\begin{itemize}
  \item monolingual baselines for English, Norwegian, and Danish,
  \item zero-shot transfer to Danish from English and Norwegian,
  \item transfer fine-tuning where a source-trained model is further fine-tuned on Danish,
  \item Danish-only supervision at multiple data budgets (10\%, 25\%, 50\%, 100\%), and
  \item linguistic-distance analysis based on WALS typological feature vectors.
\end{itemize}

The design holds the label space, evaluation metric, and Danish target data fixed, varying only the source language, encoder, and Danish supervision budget. This makes the source-language comparison interpretable without confounds.

\subsection{Models and checkpoints}

We use two multilingual encoders as the base for all token-classification models:
\begin{itemize}
  \item \textbf{mBERT}: \texttt{bert-base-multilingual-cased} \citep{devlin2019bert} — 12 layers, 110M parameters, trained on 104 languages.
  \item \textbf{XLM-R}: \texttt{xlm-roberta-base} \citep{conneau2020xlmr} — 12 layers, 125M parameters, trained on 100 languages with substantially more data than mBERT.
\end{itemize}

The English monolingual BERT baseline uses \texttt{bert-base-cased} (English-only, not multilingual) to establish a single-language ceiling. All other experiments use one of the two multilingual encoders above with a linear token-classification head for BIO tagging.

Fine-tuned transfer experiments (EN→DA FT, NO→DA FT) initialize from the best checkpoint of the corresponding zero-shot run and then continue training on Danish data. This two-stage procedure means each fine-tuned model has seen both its source-language and Danish training data.

\subsection{Training hyperparameters}

All runs share the same hyperparameter configuration unless otherwise noted:

\begin{table}[t]
\centering
\small
\begin{tabular}{ll}
\toprule
Hyperparameter & Value \\
\midrule
Learning rate          & $2\times10^{-5}$ (AdamW) \\
Train batch size       & 8 \\
Eval batch size        & 16 \\
Max sequence length    & 256 tokens \\
Training epochs        & 4 \\
Weight decay           & 0.01 \\
Random seed            & 42 \\
Early stopping         & Best checkpoint by dev F1 \\
\bottomrule
\end{tabular}
\caption{Shared hyperparameters used across all experiments. All runs used a single seed; multi-seed replication is a known limitation (see Section~\ref{sec:discussion}).}
\label{tab:hyperparams}
\end{table}

Hyperparameters were not tuned per source language; the same configuration was used for every run. This is conservative and means results could potentially improve with per-language tuning, but it also means the comparison is fair in the sense that no system received an advantage from hyperparameter search.

\subsection{Evaluation}

The primary metric is strict span-level F1 computed by \texttt{span\_f1.py}, which requires exact match on both entity boundaries and type label. We also report unlabeled span F1 (exact boundaries, any label) and loose span F1 (partial boundary overlap with matching label). These three metrics together let us distinguish missed spans from boundary errors from labeling errors.

The English test set is masked (no gold labels available), so test results are only reported for Danish and Norwegian experiments. All reported test scores use the Danish DDT test split for Danish-target experiments and the NorNE test split for Norwegian source baselines.

\subsection{Statistical testing}

Because all experiments use a single random seed, we cannot report mean $\pm$ standard deviation. To assess whether the core EN-vs-NO comparison is statistically reliable, we run a paired approximate randomisation test (paired bootstrap with 10{,}000 iterations) on the dev and test prediction files. Results are reported in Section~\ref{sec:results}.
