\section{Related Work}

Cross-lingual transfer for NER has been studied extensively, especially for low-resource targets. \citet{plank2020danish} showed that English-to-Danish transfer can already be useful in zero-shot settings and improves further with small amounts of Danish supervision --- a finding our experiments corroborate and extend to Norwegian. More generally, \citet{fu2023heterogeneous} found that source--target similarity matters, especially when languages share structural properties.

For Danish NER specifically, two resources are central to the literature. \citet{hvingelby2020dane} introduced DaNE, a manually annotated Danish NER corpus with PER, LOC, ORG, and MISC labels, which is the annotation standard our data follows. \citet{enevoldsen2021dacy} presented DaCy, a high-performance Danish NLP pipeline built on top of the DaNE standard and demonstrating what strong Danish-supervised models can achieve. Our Danish monolingual baselines are most directly comparable to results reported in these two works.

Multilingual encoders are the main enabler of zero-shot transfer. BERT \citep{devlin2019bert} introduced the transformer pretraining recipe that became the foundation for many multilingual systems. XLM-R \citep{conneau2020xlmr} improved on this by scaling data and coverage, which is why it often outperforms mBERT in cross-lingual settings. For Danish NER, the key question is how well the shared encoder space aligns source and target representations.

Linguistic distance has been used to predict and explain transfer behavior. Typological databases such as URIEL+ \citep{khan2025urielplus} and language-distance collections such as DistALS \citep{goot2025distals} provide structured comparisons across languages. \citet{sohn2024zeroshot} showed that surface similarity can help in zero-shot NER transfer even for low-resource languages. Our work extends this line by computing concrete WALS-feature distances and linking them to observed transfer gaps.

The dataset we use follows the Universal NER benchmark \citep{mayhew2024universal}, which provides multilingual NER data in consistent IOB2 format across languages. This shared setup removes annotation-format confounds from the cross-lingual comparison, making source-language effects directly interpretable.
